Propositions as Sessions, Mechanically

### 2015-04-04 paper/formalisation.tex
Freewriting to brainstorm about cut elim, GV, CPS translation.

The formalisation was performed using the Coq proof
assisstant~\cite{Coq:manual}. In the spirit of the POPLMARK
challenge~\cite{Aydemir:2005:MMM}, the development utilises available
libraries and infrastructure as much as possible. In particular,
\citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM} developed a metatheory
library for handling, among other things, association lists (environments) and
variable representation. The current development uses the version of the
library distributed by~\citeauthor{Park:2014:MMW}~\cite{Park:2014:MMW} which
has been updated for Coq version 8.4pl4. One goal of the formalisation effort
was to minimise the extent to which the development differs from the
pen-and-paper presentation and avoid questions of adequacy of the
representation. Some changes were made to ease the mechanisation and are
highlighted where they occur.

@@

@@

The typing judgements for the functional language GV are shown in
Figure~\ref{??}. Note that I have altered the $n$-ary branch and choice
constructs to the restricted binary versions in order to avoid tricky
formulations. The change is not restrictive however, since CP itself provides
only binary versions of plus ($\oplus$) and with ($\with$) constructs.

@@

explicit limiting and unlimiting of abstraction to ensure rules are syntax
directed - permits inductive argument on terms when translating from gv to cp

metatheory library and the library of tactics from software foundations are
used extensively: metatheory does not handle multiple binders (check this?)
maybe add a generic tactic to handle multiple binders... or perhaps add a
lemmas for handling binders

@@

type inference of gv terms required cluttered definition of translation with
impossible error cases. avoided a dependently typed solution due to
difficulties encountered earlier in the development. james mckinna et
al. suggest encoding the successful cases in an inductive type and then
defining the function to act on these cases. will this really work? this means
a graph of positive cases for types then a function definition for it too

translation uses the coq-ext-lib from michela for monadic notation

@@

>>>>

The formalisation was performed using the Coq proof
assistant~\cite{Coq:manual}. In the spirit of the POPLMARK
challenge~\cite{Aydemir:2005:MMM}, the development utilises available
libraries and infrastructure as much as possible. In particular,
\citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM} developed a metatheory
library for handling, among other things, association lists (environments) and
variable representation. The current development uses the version of the
library distributed by~\citeauthor{Park:2014:MMW}~\cite{Park:2014:MMW} which
has been updated for Coq version 8.4pl4. One goal of the formalisation effort
was to minimise the extent to which the development differs from the
pen-and-paper presentation and avoid questions of adequacy of the
representation. Some changes were made to ease the mechanisation and are
highlighted where they occur.

@@

% ... subject reduction....

subject reduction proof proceeds by induction on the reduction relation then
is automatic by appeal to the proved well-typed reductions.

% ... cut elimination ....

more elaborate than first thought. complicated measure.  as it stands the
current definitions of well-typed terms and reduction are not amenable to the
proving of the cut elimination theorem. some method of measure is necessary in
order to develop some notion of termination and also some relationship to the
the the underlying subprocesses additionally there is also the issue of the
representation which includes cases that are not immediately clear how to
prove for instance the commuting conversions are one such example the
reduction is only defined if the logical operator corresponds to a free
variable but induction on the typing judgement leads to two cases one where
the logical operator is applied to the top-level cut variable it is not
immediately clear how to eliminate this from happening a relationship between
the two subprocesses of the cut would need to be stated to prevent them from
differing so that a principal cut reduction rule may be applied yes only then
may the the rule be reduced to a cut-free thing if we had a relationship
between them then there would be some kind of thing that would allow
destruction on one then fixing the other in the case of a bvar equal to zero
\todo{how to proceed?}

@@

The typing judgements for the functional language GV are shown in
Figure~\ref{??}. Note that I have altered the $n$-ary branch and choice
constructs to the restricted binary versions in order to avoid tricky
formulations. The change is not restrictive however, since CP itself provides
only binary versions of plus ($\oplus$) and with ($\with$)
constructs. Explicit limiting and unlimiting of abstraction was chosen to
enable the inclusion of these rules in the translation from GV to CP, which is
performed by induction on GV terms.

@@

similar infrastructure for gv required as for cp copy and paste job

@@

require type inference on gv terms to construct corresponding cp terms but we
are assured that this is okay since the theorem only provides a well-typed gv
term in a context that is well-defined

type inference of gv terms required cluttered definition of translation with
impossible error cases. avoided a dependently typed solution due to
difficulties encountered earlier in the development. james mckinna et
al. suggest encoding the successful cases in an inductive type and then
defining the function to act on these cases. will this really work? this means
a graph of positive cases for types then a function definition for it too this
seems rather pathetic because i have to define th einference on terms and it
seems wrong i have then to handle environments but also nonenvironments i
can't remember what they are called unbound variables that is it they have to
be handled as well so we have all this clutter and the types too that is the
important part but the environments then have to be handled linearly
unfortunately very bad because we could just use the well typed term relation
for most of this headachey breaky stuff i'm not sure about that but what is
can i say so for example lets see see no i also have to thing about the
general guidelines i can be giving for this section so we have something to
say about guidelines do we yes of course we have can say that out of this we
realised that it is important for a representation to be good enough to
succintly express that which we wish to say in particular the graph relation
the inductive type that we define propositionally is better one would think
than the other way the monadic possible failure way is not producing a valid
thing that we can inductively induct on so we concern ourself with the proof
of translation preserving the typed term and i think it is important to
remember the problem we had with the inductive hypothesis we had something do
with translating the inner term of a weakening term in a larger context than
the one we had available and that was to do with the destruction of the term
inference thing i think no no it was something else we had the inner term and
we had it well typed in an environment that was smaller sure and then our
thing our function was saying we had some environment that typed both the
inner term and the outer weakening term but we know that is wrong so wrong so
here is perhaps where the inductive type graph relation thing wins out in that
we can have a hypothesis for the inner term which is sufficiently small for
the environment and we pattern match not only on the term but also on the
environment yes that is we will do because it seems to work only when we get
to the messy two the messy two constructor caases will this become a bigg
problem inthat we will then have all this stuff about environment being split
and needing to be recombined in the composite process that is a term not a
process but the point still stands it will need to be about the term and the
splitting is really a duplication of the unfortunately the well typed term
relation but otherwise i cannot think how to reuse or obtain that information
for the translation to cp

Translation uses the coq-ext-lib from michela for monadic notation

@@

\todo{how might we make progress on this?}

\subsection{Issues and Guidelines}

modularised development; tactics slow prover took a long time to reduce proofs
so split the structure of the development \todo{picture of development
  structure}.

tactic performance is an issue to be addressed due to the slow proof search a
future work (things like deciding equations on finite sets)

One of the design goals of the development was to re-use existing libraries
wherever possible. two such instances proved extremely useful. first is the
Metatheory library developed at upenn. second, also from upenn, is the library
of tactics from software foundations.

software foundations helped in many ways: simplifying tacticals, ease of
introducing proved assertions, etc. limited use of names introduced by prover
leading to more robust development.

used extensively: metatheory does not handle multiple binders (check this?)
maybe add a generic tactic to handle multiple binders... or perhaps add a
lemmas for handling binders

Linearity create issues with typing some examples; the splitting of the
environment was non-deterministic so required side-conditions to enforce a
certain splitting of binders. For example the rule for associativity of
cut. \todo{how to handle? is there a better way?}

Base rules and rules with binders required permutation assumptions to allow
any ordering of the constituent contexts. a lot of the proofs for well-typed
reductions involved small amounts of permuting environments which were largely
similar across the proofs. while some simple lemmas and tactics where created
to handle the most tedious elements these did not make effective use of
hypotheses from the context. Improved proof search for permutations is
required to simplify such developments.
